---
title: "CORD19 dates"
author: "Bianca Kramer, for ASReview"
date: "March 30, 2020"
output:
  pdf_document: default
  html_document: default
---

Retrieved dates for DOIs ('date created') and PMCID ('date first published') for records in CORD19 dataset that do not have date in proper format

CORD19 retrieved from: [https://ai2-semanticscholar-cord-19.s3-us-west-2.amazonaws.com/2020-03-27/metadata.csv](https://ai2-semanticscholar-cord-19.s3-us-west-2.amazonaws.com/2020-03-27/metadata.csv)  
Version of CORD19 used: 20200327 (v5)  
Date of sampling: 20200329

Scripts and data in:
[CORD19v5_R](/CORD19v5_R) [_fix path display_]

Results files:
[CORD19v5_R/output/CORD19id_date.csv](/CORD19v5_R/output/CORD19id_date.csv)  [_fix path display_]  
(source, publication ids, date)

[CORD19v5_R/output/CORD19id_date.csv](/CORD19v5_R/output/CORD19id_date.csv)  [_fix path display_]  
(source, publication ids, date)

## Workflow

0. **Get ids from CORD19 dataset**
    + download CORD19 [n = 45,774] 
      + (NB specify date format for publish_time, this will only retrieve dates that are in standard format)
    + keep columns for source, doi, pmcid, pmid, publish_time [n = 45,774]
    + check duplicate* [n = 45,774] <- no duplicate records
1. **For records without proper date format that have DOIs, get date created from Crossref API**
    + select records without date in proper format [n = 1522]
    + of these, select records with DOIs [n = 971]
    + retrieve data from Crossref [n = 781]
      + (NB notation of dois from CORD19 v4 onwards is harmonized)
    + match dates back to CORD19 ID data 
      + (NB for match, use lowercase for both sets of dois)
2. **For remaining records with PMCID, get date first published from EuropePMC API**
    + select remaining records without date in proper format* [n = 740]
    + of these, select records with PMCID [n = 255]
    + retrieve data from EuropePMC [n = 255]
    + match dates back to CORD19 ID data
3. **Analyze remaining records**
    + select remaining records without date in proper format* [n = 485]
    + of these, select records with PMID [n= 165] <- do NOT pursue at this time)
4. **Create CORD-19 subset by date**
    + + download CORD19 [n = 45,774] 
      + (NB specify character format for publish_time, this will retrieve dates that are in standard format)
    + copy column 'publish_time' into column 'date' as date format, (this will only keep dates that are in standard format)
    + match retrieved dates by doi, pmcid and pmid
    + filter on records with date >= 2019-12-01, or publish_time == 2020 [n = 4001]

*This includes records without DOI, and records where DOI data was not retrieved from Crossref

## Results

Load libraries
```{r message=FALSE, warning=FALSE}
#install.packages("tidyverse")
#install.packages("lubridate")
library(tidyverse)
library(lubridate)
```

Read file with matched dates
```{r}
CORD19id_date <- read_csv(
  "CORD19v5_R/output/CORD19id_date.csv",
  col_types = cols(pmcid = col_character(),
                   pubmed_id = col_character())
  )
```



What is the distribution (by year) of records with date information?
```{r}

#create column for year
CORD19id_date_year <- CORD19id_date %>%
  mutate(year = lubridate::year(date))

#quick plot
date_plot <- CORD19id_date_year %>%
  filter(!is.na(year))

ggplot(date_plot, aes(year)) +
  geom_histogram(binwidth = 1, fill="steelblue") + 
  theme_minimal()
                 
```


What is the average number of articles per month for each year?  

```{r}
ave_month <- CORD19id_date_year %>%
  filter(!is.na(date)) %>%
  group_by(year) %>%
  count() %>%
  mutate(month = case_when(
    year != 2020 ~ round((n/12)),
    year == 2020 ~ round((n/3)))
    )

ggplot(data=ave_month, aes(x=year, y=month)) +
  geom_bar(stat="identity", width=0.8, fill="steelblue") +
  theme_minimal() +
  labs(y = "articles / month")


```



## Outstanding issues:
1. Validate results against years in original dataset

2. Pretty pictures!
    + also plot by month for 2019-2020



  
